Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-01-28 paper/1_Initial/manuscript2.qmd
updated manuscript

---
title: "Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates"
shorttitle: "IAS Meta-analysis"
author:
  - name: Ana Neves
    orcid: 0009-0006-0020-7599
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
# author-note:
#   status-changes: 
#     affiliation-change: null
#     deceased: null
#   disclosures:
#     study-registration: null
#     data-sharing: null
#     related-report: null
#     conflict-of-interest: null
#     financial-support: null
#     gratitude: null
#     authorship-agreements: null
    #   # - Project administration
    #   # - Data curation
    #   # - Formal Analysis
    #   # - Investigation
    #   # - Visualization
    #   # - Writing – original draft
    #   # - Writing – review & editing
    # Roles are optional. 
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    #   - Conceptualization
    #   - Data curation
    #   - formal Analysis
    #   - Funding acquisition
    #   - Investigation
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing – original draft
  #   #   - Writing – review & editing
  # - name: Magdalena Pfaff
  #   orcid: 0009-0006-2386-7936
  #   affiliations:
  #     - ref: "id1" 
  # - name: Dominique Makowski
  #   corresponding: true
  #   orcid: 0000-0001-5375-9967
  #   email: D.Makowski@sussex.ac.uk
  #   url: https://realitybending.github.io/
  #   roles:
  #     - Conceptualization
  #     - Data curation
  #     - Formal Analysis
  #     - Funding acquisition
  #     - Investigation
  #     - Methodology
  #     - Project administration
  #     - Resources
  #     - Software
  #     - Supervision
  #     - Validation
  #     - Visualization
  #     - Writing – original draft
  #   affiliations:
  #     - ref: "id1"
  #     - name: "University of Sussex"
  #       department:  "Sussex Centre for Consciousness Science"
author-note:
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      :::
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
abstract: |
  Blabla the abstract blabla.
keywords: [keyword1, keyword2, keyword3]
floatsintext: true
numbered-lines: true
bibliography: bibliography.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    include-in-header:
      - text: |
          \usepackage{lscape}
  apaquarto-docx: default
  apaquarto-html: default
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: console
---

@@

The use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical. However, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information. These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively. This approach offers useful and interesting measures [@murphy2019; @lin2023].

@@

<!-- lack of convergence between self-reported questionnaires -->

A recent systematic review by @desmedt2022measures examined whether various questionnaires designed to measure "interoceptive sensibility" truly assess the same construct. Notably, @garfinkel2015knowing defines interoceptive sensibility as the self-reported tendency to focus on and detect internal sensations, whereas @khalsa2018interoception defines it more narrowly, excluding the ability to detect these signals. According to @desmedt2022measures, most authors adopt the latter definition. The review found that these questionnaires measure related but distinct constructs, leading researchers to commit the 'jingle fallacy'—mistakenly treating them as equivalent measures of interoceptive sensibility.

**TODO: do a better job**

<!-- Moreover, the current subjective measures do not correlate well with the objective measures [@murphy2019]. -->

Few studies have examined the correlation between objective interoceptive measures, such as the Heartbeat Counting Task (HCT; Schandry, 1981) and the Heartbeat Detection Task (HDT; Kleckner et al., 2015). Existing findings report weak correlations for the HCT (Murphy et al., 2019), no correlations except for the sensitivity variable of the HCT (Brand et al., 2023), and small correlations for the HDT (Brand et al., 2023). These results suggest that the IAS, a subjective measure, does not strongly align with objective interoceptive assessments.

A better understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception.

@@

<!-- was the second reference asked for the MAIA, the MAIA-2 ? -->

@@

<!-- [@morin2016] is a methodological paper on ESEM -->

@@

<!-- Relationship with other interoception measures --> 
<!-- The MAIA was introduced above -->

>>>>

---
title: "Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates"
shorttitle: "IAS Meta-analysis"
author:
  - name: Ana Neves
    orcid: 0009-0006-0020-7599
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
# author-note:
#   status-changes: 
#     affiliation-change: null
#     deceased: null
#   disclosures:
#     study-registration: null
#     data-sharing: null
#     related-report: null
#     conflict-of-interest: null
#     financial-support: null
#     gratitude: null
#     authorship-agreements: null
    #   # - Project administration
    #   # - Data curation
    #   # - Formal Analysis
    #   # - Investigation
    #   # - Visualization
    #   # - Writing – original draft
    #   # - Writing – review & editing
    # Roles are optional. 
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    #   - Conceptualization
    #   - Data curation
    #   - formal Analysis
    #   - Funding acquisition
    #   - Investigation
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing – original draft
  # - name: Magdalena Pfaff
  #   orcid: 0009-0006-2386-7936
  #   affiliations:
  #     - ref: "id1" 
  # - name: Dominique Makowski
  #   corresponding: true
  #   orcid: 0000-0001-5375-9967
  #   email: D.Makowski@sussex.ac.uk
  #   url: https://realitybending.github.io/
  #   roles:
  #   affiliations:
  #     - ref: "id1"
  #     - name: "University of Sussex"
  #       department:  "Sussex Centre for Consciousness Science"
author-note:
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      :::
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
abstract: |
  Blabla the abstract blabla.
keywords: [keyword1, keyword2, keyword3]
floatsintext: true
numbered-lines: true
bibliography: bibliography.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    include-in-header:
      - text: |
          \usepackage{lscape}
  apaquarto-docx: default
  apaquarto-html: default
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: console
---

@@

The use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical. However, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information. These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively [@murphy2019; @khalsa2018interoception]. This approach offers useful and interesting measures [@murphy2019; @lin2023].

@@

<!-- A better understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception. -->

<!-- WHATS THE ARGUMENT ?? -->
<!--   1. lack of relationship between different explicit measures of Interoception?  -->
<!--   2. lack of convergence bettween different self-report measures of interoception? -->
<!--   3. lack of correlation between subjective vs objective accuracy? -->

@@

@@

**note to dom**:@morin2016] is a methodological paper on ESEM

@@

<!-- Relationship with other interoception measures --> 
**note to dom**: the maia was introduced above

### 2025-01-28 paper/1_Initial/manuscript2.tex
updated manuscript

\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, left=2mm, colback=white, arc=.35mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

@@

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut (\citeproc{ref-khalsa2018}{Khalsa et al.,
2018}). While recent research emphasizes a key role of interoception in
a variety of processes (e.g., emotion regulation, decision making) and
of outcomes (physical and psychological well being), the field remains
clouded by concerns about how interoception is assessed.

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively. This approach offers useful and interesting
measures (\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

@@

A recent systematic review by Desmedt et al.
(\citeproc{ref-desmedt2022measures}{2022}) examined whether various
questionnaires designed to measure ``interoceptive sensibility'' truly
assess the same construct. Notably, Garfinkel et al.
(\citeproc{ref-garfinkel2015knowing}{2015}) defines interoceptive
sensibility as the self-reported tendency to focus on and detect
internal sensations, whereas
(\citeproc{ref-khalsa2018interoception}{\textbf{khalsa2018interoception?}})
defines it more narrowly, excluding the ability to detect these signals.
According to Desmedt et al. (\citeproc{ref-desmedt2022measures}{2022}),
most authors adopt the latter definition. The review found that these
questionnaires measure related but distinct constructs, leading
researchers to commit the `jingle fallacy'---mistakenly treating them as
equivalent measures of interoceptive sensibility.

\textbf{TODO: do a better job}

Few studies have examined the correlation between objective
interoceptive measures, such as the Heartbeat Counting Task (HCT;
Schandry, 1981) and the Heartbeat Detection Task (HDT; Kleckner et al.,
2015). Existing findings report weak correlations for the HCT (Murphy et
al., 2019), no correlations except for the sensitivity variable of the
HCT (Brand et al., 2023), and small correlations for the HDT (Brand et
al., 2023). These results suggest that the IAS, a subjective measure,
does not strongly align with objective interoceptive assessments.

A better understanding of what is being measured with different
questionnaires and dimensions, as well as their potential overlaps with
other constructs (e.g., alexithymia, body awareness), is thus needed to
clarify the role of self-reports in the assessment of interoception.

@@

@@

@@

\bibitem[\citeproctext]{ref-garfinkel2015knowing}
Garfinkel, S. N., Seth, A. K., Barrett, A. B., Suzuki, K., \& Critchley,
H. D. (2015). Knowing your own heart: Distinguishing interoceptive
accuracy from interoceptive awareness. \emph{Biological Psychology},
\emph{104}, 65--74.

@@

@@

>>>>

\par{     \begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, toprule=.15mm, arc=.35mm, breakable, opacityback=0, colback=white]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript2_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

@@

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut (\citeproc{ref-khalsa2018}{Khalsa, Adolphs,
Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel, Lane,
Mehling, Meuret, et al., 2018}). While recent research emphasizes a key
role of interoception in a variety of processes (e.g., emotion
regulation, decision making) and of outcomes (physical and psychological
well being), the field remains clouded by concerns about how
interoception is assessed.

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively (\citeproc{ref-khalsa2018interoception}{Khalsa,
Adolphs, Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel,
Lane, Mehling, \& others, 2018}; \citeproc{ref-murphy2019}{Murphy et
al., 2019}). This approach offers useful and interesting measures
(\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

@@

@@

\textbf{note to dom}:Morin et al. (\citeproc{ref-morin2016}{2015}){]} is
a methodological paper on ESEM

@@

\textbf{note to dom}: the maia was introduced above

@@

@@

\bibitem[\citeproctext]{ref-khalsa2018interoception}
Khalsa, S. S., Adolphs, R., Cameron, O. G., Critchley, H. D., Davenport,
P. W., Feinstein, J. S., Feusner, J. D., Garfinkel, S. N., Lane, R. D.,
Mehling, W. E., et al. (2018). Interoception and mental health: A
roadmap. \emph{Biological Psychiatry: Cognitive Neuroscience and
Neuroimaging}, \emph{3}(6), 501--513.

@@

\bibitem[\citeproctext]{ref-morin2016}
Morin, A. J. S., Arens, A. K., \& Marsh, H. W. (2015). A bifactor
exploratory structural equation modeling framework for the
identification of distinct sources of construct-relevant psychometric
multidimensionality. \emph{Structural Equation Modeling: A
Multidisciplinary Journal}, \emph{23}(1), 116--139.
\url{https://doi.org/10.1080/10705511.2014.961800}
